Carbonate chemistry at Ny-Ålesund

### 2022-08-08 awipev-CO2.Rmd
pCO2 clean up code, print fig.

# downloaded 2021-04-14
# http://ebas.nilu.no/Pages/DataSetList.aspx?key=3673A72690B343B1AEC0F1673FE8C09D
file_names <- c("atmospheric_co2_zeppelin/NO0042G.20160101000000.20180420115103.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20170101000000.20180413180000.online_crds.carbon_dioxide.air.7mo.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", 
"atmospheric_co2_zeppelin/NO0042G.20180101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20190101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20200101000000.20210413000000.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas")
atm_co2_new <- NULL
year_co2 <- ymd_hm(paste0(c(2016:2020), "-01-01 00:00"))
for (i in 1:5) {
tmp <-
  read_table(file = paste0(path, file_names[i]),
  comment = "#",
  #skip = 182,
  col_names = FALSE,
  na = c("9999.999", "99.999", "9999.990", "9999.99")
  ) %>%
  dplyr::mutate(co2_air = X3,
  datetime = year_co2[i] + X1*24*3600) %>% 
  dplyr::select(datetime, co2_air)
  tmp$datetime <- lubridate::round_date(tmp$datetime, unit="1 minutes")
atm_co2_new <- bind_rows(atm_co2_new, tmp)
  }
 # combine the two sources of CO2 data
# the new source has missing data in 2018, so the data from the old source
# are used for 2018
atm_co2 <- dplyr::full_join(atm_co2_new, atm_co2_old, by = "datetime")
atm_co2 <- atm_co2 %>% 
  dplyr::rename(co2_air_new = co2_air.x, co2_air_old = co2_air.y) %>%
  dplyr::mutate(co2_air = ifelse(year(atm_co2$datetime) == 2020, co2_air_new, co2_air_old))
# Below was an attempt to smooth but
# zz <- as_tsibble(atm_co2) %>% 
#   mutate(smoothed_co2_air = slide_dbl(co2_air, .f = mean, .before = 12, .after = 12, .complete = TRUE))
# ggplot(zz) +
#   geom_line(aes(x = datetime, y = smoothed_co2_air))
d_all <- dplyr::left_join(d_all, dplyr::select(atm_co2, datetime, co2_air), by = "datetime") %>% 
  dplyr::select(sort(current_vars()))

p_atm_co2 <- d_all %>% 
  #tidyr::pivot_longer(cols = c(pco2_air, pco2_air_old), names_to = "dataset", values_to = "pco2") %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = co2_air), col = "blue", size = 0.15, na.rm = TRUE) +
  #facet_grid(rows = vars(dataset)) +
  #geom_point(aes(x = datetime, y = pco2_air_old), col = "grey", size = 0.15, na.rm = TRUE) +
  labs(title = "Atmospheric CO2 at Zeppelin", x = "Time", y = "CO2 (ppm)") +
  Mytheme(size_labs = 8)
#p_atm_co2
ggsave(filename = "figures/atm_co2.png", plot = p_atm_co2, width = 20, units = "cm")
```

@@

# Combine plots
g1 <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/time_series.png", plot = g1, width = 15, height = 30, units = "cm")
#rm(list = ls(pattern = "ts_")) #delete big ggplot objects
#gc() #garbage collection
```

@@

# identify extremes
d_bp <- d_bp %>%
  dplyr::mutate(
    ext_sal_fb_mix = ifelse(
      sal_fb_mix < dec_sal_fb_mix[1] |
        sal_fb_mix > dec_sal_fb_mix[2],
      sal_fb_mix,
      NA
    ),
    ext_temp_insitu_11m = ifelse(
      temp_insitu_11m < dec_temp_insitu_11m[1] |
        temp_insitu_11m > dec_temp_insitu_11m[2],
      temp_insitu_11m,
      NA
    ),
    ext_pco2_fb = ifelse(pco2_fb < dec_pco2_fb[1] |
                           pco2_fb > dec_pco2_fb[2], pco2_fb, NA),
    ext_pHint_tot_sf = ifelse(
      pHint_tot_sf < dec_pHint_tot_sf[1] |
        pHint_tot_sf > dec_pHint_tot_sf[2],
      pHint_tot_sf,
      NA
    ),
    ext_at_calc = ifelse(at_calc < dec_at_calc[1] |
                           at_calc > dec_at_calc[2], at_calc, NA)
  )
ts_sal_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb_mix)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_sal_fb_mix), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Salinity (< 8 m)", y = "Salinity", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_temp_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_insitu_11m)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_temp_insitu_11m), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ temperature (11 m)", y = "Temperature (°C)", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pco2_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_fb)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pco2_fb), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=expression(paste(pCO[2], " at 11 m and in situ temperature")), y = TeX("pCO$_2$ ($\\mu atm$)"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pH_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pHint_tot_sf), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ pH below 8 m at in situ temperature", y = expression(paste(pH[T])), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_at_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_at_calc), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Total alkalinity calculated from salinity", y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
# Combine plots
g2 <- (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g2 <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_boxplots.png", plot = g2)
#rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
#gc() #garbage collection
```
```{r fig ts and bp for paper}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_ts_boxplots.png", plot = g)
rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
gc() #garbage collection
```

>>>>

# downloaded 2021-04-14
# http://ebas.nilu.no/Pages/DataSetList.aspx?key=3673A72690B343B1AEC0F1673FE8C09D
file_names <- c("atmospheric_co2_zeppelin/NO0042G.20160101000000.20180420115103.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20170101000000.20180413180000.online_crds.carbon_dioxide.air.7mo.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", 
"atmospheric_co2_zeppelin/NO0042G.20180101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20190101000000.20210413133953.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas", "atmospheric_co2_zeppelin/NO0042G.20200101000000.20210413000000.online_crds.carbon_dioxide.air.1y.1h.NO01L_picarro2_G2401.NO01L_picarro.lev2.nas")
atm_co2_new <- NULL
year_co2 <- ymd_hm(paste0(c(2016:2020), "-01-01 00:00"))
for (i in 1:5) {
tmp <-
  read_table(file = paste0(path, file_names[i]),
  comment = "#",
  #skip = 182,
  col_names = FALSE,
  na = c("9999.999", "99.999", "9999.990", "9999.99")
  ) %>%
  dplyr::mutate(co2_air = X3,
  datetime = year_co2[i] + X1*24*3600) %>% 
  dplyr::select(datetime, co2_air)
  tmp$datetime <- lubridate::round_date(tmp$datetime, unit="1 minutes")
atm_co2_new <- bind_rows(atm_co2_new, tmp)
  }
 # combine the two sources of CO2 data
# the new source has missing data in 2018, so the data from the old source
# are used for 2018
atm_co2 <- dplyr::full_join(atm_co2_new, atm_co2_old, by = "datetime")
atm_co2 <- atm_co2 %>% 
  dplyr::rename(co2_air_new = co2_air.x, co2_air_old = co2_air.y) %>%
  dplyr::mutate(co2_air = ifelse(year(atm_co2$datetime) == 2020, co2_air_new, co2_air_old))

d_all <- dplyr::left_join(d_all, dplyr::select(atm_co2, datetime, co2_air), by = "datetime") %>% 
  dplyr::select(sort(current_vars()))

p_atm_co2 <- d_all %>% 
  #tidyr::pivot_longer(cols = c(pco2_air, pco2_air_old), names_to = "dataset", values_to = "pco2") %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = co2_air), col = "blue", size = 0.15, na.rm = TRUE) +
  labs(title = "Atmospheric CO2 at Zeppelin", x = "Time", y = "CO2 (ppm)") +
  Mytheme(size_labs = 8)
p_atm_co2
ggsave(filename = "figures/atm_co2.png", plot = p_atm_co2, width = 20, units = "cm")
```

@@

# Combine plots
g1 <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at)
g1 <- g1 + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/time_series.png", plot = g1, width = 15, height = 30, units = "cm")
#rm(list = ls(pattern = "ts_")) #delete big ggplot objects
#gc() #garbage collection
```

@@

# identify extremes
d_bp <- d_bp %>%
  dplyr::mutate(
    ext_sal_fb_mix = ifelse(
      sal_fb_mix < dec_sal_fb_mix[1] |
        sal_fb_mix > dec_sal_fb_mix[2],
      sal_fb_mix,
      NA
    ),
    ext_temp_insitu_11m = ifelse(
      temp_insitu_11m < dec_temp_insitu_11m[1] |
        temp_insitu_11m > dec_temp_insitu_11m[2],
      temp_insitu_11m,
      NA
    ),
    ext_pco2_fb = ifelse(pco2_fb < dec_pco2_fb[1] |
                           pco2_fb > dec_pco2_fb[2], pco2_fb, NA),
    ext_pHint_tot_sf = ifelse(
      pHint_tot_sf < dec_pHint_tot_sf[1] |
        pHint_tot_sf > dec_pHint_tot_sf[2],
      pHint_tot_sf,
      NA
    ),
    ext_at_calc = ifelse(at_calc < dec_at_calc[1] |
                           at_calc > dec_at_calc[2], at_calc, NA)
  )
ts_sal_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb_mix)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_sal_fb_mix), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Salinity (< 8 m)", y = "Salinity", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_temp_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_insitu_11m)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_temp_insitu_11m), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ temperature (11 m)", y = "Temperature (°C)", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pco2_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_fb)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pco2_fb), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=expression(paste(pCO[2], " at 11 m and in situ temperature")), y = TeX("pCO$_2$ ($\\mu atm$)"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pH_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pHint_tot_sf), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ pH below 8 m at in situ temperature", y = expression(paste(pH[T])), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_at_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_at_calc), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Total alkalinity calculated from salinity", y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
# Combine plots
g2 <- (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g2 <- g2 + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_boxplots.png", plot = g2)
#rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
#gc() #garbage collection
```
```{r fig ts and bp for paper}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_ts_boxplots.png", plot = g)
rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
gc() #garbage collection
```
